Reproducible Research 
==========================
# Peer Assessment 1

Make a R markdown file to show the process of data analysis on the personal movement data collected from monitoring devices.

## 1 Loadding and preprocessing the data
````{r ,echo=TRUE,results='hide'}
#prepare the libraries we need
library(knitr)
library(ggplot2)

# load data
setwd("/Develop/R/Coursera/ReproResrch")
data=read.csv("activity.csv")
```
-------------


##  2  What is mean total number of steps taken per day
````{r ,echo=TRUE ,fig.width=12}
#Get the sum of steps taken for each date and turn the reusslt to dataframe
sum.numofStepsPerDay <- tapply(data$step, data$date, sum,na.rm=TRUE)
sum.df=data.frame(date=names(sum.numofStepsPerDay),sumStepsPerDay=sum.numofStepsPerDay)
# Make Histgram using ggplot
ggplot(sum.df, aes(x=date,y=sumStepsPerDay)) + geom_histogram(binwidth=.5,stat="identity")
```

````{r ,echo=TRUE}
#The mean and median total number if steps taken per day
mean.StepsPerDay= mean(sum.df$sumStepsPerDay)
median.StepsPerDay=median(sum.df$sumStepsPerDay)
```
### [ Answer ] : The mean is `r mean.StepsPerDay` and median is `r median.StepsPerDay` 

-------------


##  3  What is the average daily activity pattern?
````{r ,echo=TRUE,fig.width=12}
#Get the average steps taken for each interval and turn the reusslt to dataframe
avg.avgStepsPerInterval <- tapply(data$step, data$interval, mean,na.rm=TRUE)
avg.df=data.frame(interval=names(avg.avgStepsPerInterval),avgSteps=avg.avgStepsPerInterval)
avg.df$interval=as.numeric(as.character(avg.df$interval))

#Time series plot of the average number of steps taken (averaged across all days) versus the 5-minute intervals
ggplot(avg.df,aes(x=interval,y=avgSteps))+geom_line()
#Find the interval with max avgSteps taken
maxAvgStepsPerInterval=avg.df[which.max(avg.df$avgSteps),]$interval
print(maxAvgStepsPerInterval)
```
### [ Answer ] : The interval with max value of steps accross all days : `r maxAvgStepsPerInterval` 

-------------


##  4 Imputing missing values
### 4.1 Total number of rows with NAs
````{r ,echo=TRUE}
#Caculate the total number of rows with NAs
numOfNa=sum(is.na(data$steps))
```
#### Report of the total number of rows with NAs : `r numOfNa`

## 4.2 Derive a Strategy to fill the NAs
### [ Answer ] : The strategy : Fill the NAs with average steps for that interval
````{r }
#
data2=data
for (i in 1:nrow(data2)){
    # if steps is NA
    if( is.na(data2[i,]$steps)) {
      # fill the NA with the avgSteps of that i
      tarInterval<-data2[i,]$interval # the interval
      data2[i,]$steps<-avg.df[avg.df$interval==tarInterval,]$avgSteps
    }
  }
#Verify if the data2 is without NA
sum(is.na(data2$steps))
```
### 4.3 Create a new dataset that is equal to the original dataset but with the missing data filled in
````{r }
#Verify if the data2 is without NA
sum(is.na(data2$steps))
```
### 4.4 Make a histogram of the total number of steps taken each day (Without NAs!!)
````{r echo=TRUE,fig.width=12}
#Caculate the total number of steps taken for each Date (using data2)
sum2.numofStepsPerDay <- tapply(data2$step, data$date, sum)
sum2.df=data.frame(date=names(sum2.numofStepsPerDay),sumStepsPerDay=sum2.numofStepsPerDay)
#Histgram using ggplot
ggplot(sum2.df, aes(x=date,y=sumStepsPerDay)) + geom_histogram(binwidth=.5,stat="identity")

# Caculate the mean and median
mean2.StepsPerDay= mean(sum2.df$sumStepsPerDay)
median2.StepsPerDay=median(sum2.df$sumStepsPerDay)
```
### [ Answer ]  : The data (has NAs) mean is `r mean.StepsPerDay` and median is `r median.StepsPerDay`
### [ Answer ]  : The data2 (has no NAs) mean is `r mean2.StepsPerDay` and median is `r median2.StepsPerDay`

--------------


## 5 Are there differences in activity patterns between weekdays and weekends?
###   5.1 Create a new factor variable in the dataset with two levels 
````{r }
# useing data without NAs: data2 to get the weekday character Vector
weekdayVtr<-weekdays(as.Date(data2$date))
# check every elements are weekend or weekday (p.s. I use the Chinese words meanning Sunday and Saturday because the seeting )
weekdayFactor<-unlist(lapply(weekdayVtr, function(x) if (x=="周日" |x=="周六") {x<-"weekend" } else{ x<-"weekday"} ))
# transform vector as factor, and put  back to data2
data2<-cbind(data2,as.factor(weekdayFactor))

avg2.avgStepsPerInterval <- tapply(data2$step, data2$interval, mean)
avg2.df=data.frame(interval=names(avg2.avgStepsPerInterval),avgSteps=avg2.avgStepsPerInterval)
avg2.df$interval=as.numeric(as.character(avg2.df$interval))
```

###   5.2 Create a new factor variable in the dataset with two levels 
````{r }
# useing data without NAs: data2 to get the weekday character Vector
weekdayRecord=subset(data2,weekdayFactor=="weekday")
avgStepsInterval.weekday<- tapply(weekdayRecord$step, weekdayRecord$interval, mean)
avgStepsDf.weekday=data.frame(interval=names(avgStepsInterval.weekday),avgSteps=avgStepsInterval.weekday)
avgStepsDf.weekday$interval=as.numeric(as.character(avgStepsDf.weekday$interval))
avgStepsDf.weekday$weekdayFactor<-c("weekday")
avgStepsDf.weekday$weekdayFactor<-as.factor(avgStepsDf.weekday$weekdayFactor)


# Prepare weekend part for caculating average steps per interval 
weekendRecord=subset(data2,weekdayFactor=="weekend")
avgStepsInterval.weekdend<- tapply(weekendRecord$step, weekendRecord$interval, mean)
avgStepsDf.weekend=data.frame(interval=names(avgStepsInterval.weekdend),avgSteps=avgStepsInterval.weekdend)
avgStepsDf.weekend$interval=as.numeric(as.character(avgStepsDf.weekend$interval))
avgStepsDf.weekend$weekdayFactor<-c("weekend")
avgStepsDf.weekend$weekdayFactor<-as.factor(avgStepsDf.weekend$weekdayFactor)

# Combine weekend and weekday record to one dataframe
avgStepsDF.combined=rbind(avgStepsDf.weekend,avgStepsDf.weekday)

# Make plot with lattice packages:
library(lattice)
xyplot(avgSteps~interval | avgStepsDF.combined$weekdayFactor,avgStepsDF.combined,layout=c(1,2), type='l')

```
